---
title: "Weather Prediction"
author: "Olumide Karim-Mohammed"
date: "26 February 2019"
output: ioslides_presentation
---

<style>
pre{

font-size: 20px;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

```{r message = FALSE, warning=FALSE, comment=FALSE}
library(kableExtra)
library(dplyr)
library(magrittr)
library(knitr)
library(corrplot)
library(pROC)
library(tidyr)
library(ggplot2)
library(data.table)
library(caret)
library(gridExtra)
library(grid)
library(scales)
library(ggthemes)
library(rpart)
library(rpart.plot)

set.seed(123)


ConfusionMatrixInfo <- function( data, predict, actual, cutoff )
  
{          
  
  # extract the column ;
  
  # relevel making 1 appears on the more commonly seen position in
  
  # a two by two confusion matrix     
  
  predict <- data[[predict]]
  
  actual  <- relevel( as.factor( data[[actual]] ), "High")
  
  result <- data.table( actual = actual, predict = predict )
  
  # caculating each pred falls into which category for the confusion matrix
  
  result[ , type := ifelse( predict >= cutoff & actual == "High", "TP",
                            
                            ifelse( predict >= cutoff & actual == "Low", "FP",
                                    
                                    ifelse( predict <  cutoff & actual == "High", "FN", "TN" ) ) ) %>% as.factor() ]
  
  
  # jittering : can spread the points along the x axis
  
  plot <- ggplot( result, aes( actual, predict, color = type ) ) +
    
    geom_violin( fill = "grey", color = NA ) +
    
    geom_jitter( shape = 1 ) +
    
    geom_hline( yintercept = cutoff, color = "blue", alpha = 0.6 ) +
    
    scale_y_continuous( limits = c( 0, 1 ) ) +
    
    scale_color_discrete( breaks = c( "TP", "FN", "FP", "TN" ) ) + # ordering of the legend
    
    guides( col = guide_legend( nrow = 2 ) ) + # adjust the legend to have two rows 
    
    ggtitle( sprintf( "Confusion Matrix with Cutoff at %.2f", cutoff ) )+
    
    theme_economist()
  
  return(list( data = result, plot = plot ))
  
}


```

```{r}
#weather<-read.csv('C:\\Users\\7222115\\Desktop\\DS_sensor_weather.csv',header=TRUE,sep=',')
weather<-read.csv('~/Downloads/DS_sensor_weather.csv',header=TRUE,sep=',')
```

## The Question

Company X is looking to grow their IoT capabilities. One of thier systems generates 9 morning sensor readings which they would like to use to predict relative humidity in the afternoon:

*   **Ten columns including actual relative humidity in the afternoon**
*   **1095 records**


```{r echo = FALSE}
head(weather[,-11]) %>% kable() %>% 
  kable_styling(font_size = 11, full_width = FALSE) %>% 
  row_spec(0,font_size = 8,color= 'red')

```
## ..more on the data
Note left skew of ***rain_accumulation*** and ***rain_duration*** both peaking at 0


```{r echo = FALSE}

weather_nna<-weather %>% na.omit()


weather_nna_nr<-weather_nna[,-11]


weather_nna_nr_rf<-(weather_nna_nr %>% mutate(relative_humidity_pm_rf=relative_humidity_pm>30))[,-10]


weather_nna_nr_rf$relative_humidity_pm_r2<-factor(weather_nna_nr_rf$relative_humidity_pm_rf, levels = c(TRUE,FALSE), labels = c("High","Low"))

weather_nna_nr_rf$relative_humidity_pm_r2 <- relevel(weather_nna_nr_rf$relative_humidity_pm_r2, ref = "Low")



weather_cln<-weather_nna_nr_rf[,-10]

ggplot(gather(weather_cln[,-10]), aes(value)) +
  
  geom_histogram(bins = 15) +
  
  facet_wrap(~key, scales = 'free_x')

```

## ..still on the data
* omitting incomplete records we get ***1064*** obs.
* assuming relative humidity measurements over 30 to be classed as "High" we get the following proportions on ***relative_humidity_pm***:
```{r echo = FALSE}

as.data.table(prop.table(table(weather_cln$relative_humidity_pm_r2)))%>% 
  kable() %>% kable_styling(font_size = 12, full_width = FALSE)

```
* data appears fairly balanced - no need to augment sampling process
* situation might be different at other times in the year - seasons

## ..finally on data

we look at how the sensor readings interact with each other

```{r}
corrplot(cor(weather_nna_nr))
```

# Model Development


## Method 1 - ***Logistic Regression***

- Using train/test split ratio of 75:25

```{r message = FALSE, warning=FALSE, fig.align='left',fig.show='hold'}
smp_size <- floor(0.75 * nrow(weather_cln))

train_ind <- sample(seq_len(nrow(weather_cln)), size = smp_size)

train <- weather_cln[train_ind, ]

test <- weather_cln[-train_ind, ]

model_noscaling<-glm(relative_humidity_pm_r2~.,train,family='binomial')

train$pred<-predict(model_noscaling,newdata = train,type="response")

test$pred<-predict(model_noscaling,newdata = test,type="response")

conf_mat<-confusionMatrix(as.factor(ifelse(train$pred>=0.5,'High','Low')),train$relative_humidity_pm_r2)

summary_model<-summary(model_noscaling)

summary_model$coefficient %>% kable() %>% kable_styling(font_size = 8, full_width = FALSE,position = 'float_left') %>% 
  row_spec(0,font_size = 8,color= 'red')
      
```
```{r fig.align='center', fig.height=4, fig.width=4,warning=FALSE}
ggplot(test, aes( pred, color = as.factor(relative_humidity_pm_r2) ) ) +
  
  geom_density( size = 1 ) +
  
  ggtitle( "Test Set's Predicted Score" )+
  
  scale_color_economist( name = "data", labels = c( "Low", "High" ) ) +
  
  theme_economist()

conf_mat<-confusionMatrix(as.factor(ifelse(test$pred>=0.5,'High','Low')),test$relative_humidity_pm_r2)

R2<-round(1 - ( summary_model$deviance / summary_model$null.deviance ),2)*100
acc<-round(mean(test$relative_humidity_pm_r2== ifelse(test$pred>0.46,'High','Low')),2)*100

```
* with an accuracy of `r acc`% and r.sq value of `r R2`%

## ..method 1.. managing risk


```{r warning=FALSE,fig.align='center', fig.height=4, fig.width=6,fig.show='hold'}


ConfusionMatrixInfo( data = test, predict = "pred", 
                     actual = "relative_humidity_pm_r2", cutoff = .46 )$plot

```

```{r}

conf_mat$table

```
## roc

```{r}
ROC <- roc(test$relative_humidity_pm_r2, test$pred)
plot(ROC)
auc(ROC)

```

# Method 2 -Decision Trees


## The Model
- Using same dataset and train/test split ratio

```{r warning=FALSE}

weather_cln<-weather_nna_nr_rf[,-10]

train_ind <- sample(seq_len(nrow(weather_cln)), size = smp_size)

train <- weather_cln[train_ind, ]

test <- weather_cln[-train_ind, ]

weather_mod<-rpart(relative_humidity_pm_r2~.,train,method='class',control= rpart.control(cp=0))

pred<-predict(weather_mod,test,type='class')

conf_mat_dt<-confusionMatrix(pred,test$relative_humidity_pm_r2)

rpart.plot(weather_mod, type = 3, box.palette = c("green","red"), fallen.leaves = TRUE)

```

## ..more parameters{.column-2}
```{r message = FALSE, warning=FALSE, fig.height=3,fig.width=4,fig.show='hold'}
conf_mat_dt$table
ROC <- roc(ifelse(test$relative_humidity_pm_r2=="High",1,0),ifelse(pred=="High",1,0))
auc(ROC)
plot(ROC)
acc_dt<-round(mean(test$relative_humidity_pm_r2==pred),2)*100


```
- model accuracy is `r acc_dt`%

## Recommendations

- more data - more of the same(historical and future)
- and additional  e.g. time of year, geography etc
- specify a risk policy/ strategy with regards to the system in question
- deploy both models if possible. if you must choose opt for decision tree
- a batch training regime is recommended as the data is not too large neither is it fast and continous
- model is to be monitored regularly to spot any signs of degradation

